% !TEX root = ../../main.tex

% ----------------------------------
% labels: \label{mil4:[type]:[name]}
% ----------------------------------
% FUTURE TENSE


Quantum fluctuations from the early universe generates inhomogeneity in the form of (energy) density enhancements and deficits. The rapid expansion during inflation stretches these quantum fluctuations to cosmic scales. From this, local fluctuations in the gravitational potential are implied. The primordial plasma compresses in potential wells found in regions of high density. Analogously, the tightly coupled plasma rarefies in potential hills created in regions of low density. These mechanisms generate pressure waves propagating at the speed of sound, eventually imprinting a pattern of sound in the CMB temperature. Unless otherwise stated, compressions and decompressions in a fluid are seen from the point of view of the potential \emph{well}.

Where the fluid is compressed in a well and thus heated, there is a region \emph{nearby}\footnote{Near enough to be in causal contact with the aforementioned density enhancement.} where the fluid expands on a hill and cools down. It is convenient to decompose the fluctuations into plane wave-modes that evolve independently, as we did in~\cref{sec:mil3}. Now, if such a mode corresponds to a wavelength that is shorter than the particle horizon, \emph{and} the photons are tightly coupled to the baryons, the temperature fluctuation will oscillate in time as the fluid compresses and rarefies. At recombination, sound waves stop oscillating and are frozen in. The last scattering surface corresponds to a maximal length for these sound waves to have travelled; the sound horizon at decoupling, briefly mentioned in~\cref{sec:mil2}. There is then a mode ($k_1$) that had just enough time to compress once, necessarily with half a wavelength equal this distance. The mode ($k_2$) with this distance as its whole wavelength, will have completed one cycle of compression/decompression at decoupling. These modes (2\textpi~over comoving wavelength) are then $k_n=n\pi/s_*$, where $n\in \mathbb{N}$ by extension.

Now, the perturbations from~\cref{sec:mil3} can be combined with an estimate of the post-inflation dilated quantum fluctuations to form theoretical power spectra, in particular those of the photon temperature and non-relativistic matter today. We will in this section produce these observables using the simplest concordance model of inflation. 

\subsection{Theory}\label[sec]{mil4:sec:theo}
\input{sections/milestone4/theory.tex}

\subsection{Implementation details}\label[sec]{mil4:sec:imp}
\input{sections/milestone4/implementation.tex}

\subsection{Results}\label[sec]{mil4:sec:res}
\input{sections/milestone4/results.tex}

\subsection{Discussion}\label[sec]{mil4:sec:disc}
\input{sections/milestone4/discussion.tex}