% !TEX root = ../../main.tex

% ----------------------------------
% labels: \label{mil1:[type]:[name]}
% ----------------------------------
% FUTURE TENSE


The first ingredient in the aforementioned program is a numerical framework describing the (unperturbed) background geometry. Said geometry is determined by the Friedmann-Robertson-Walker (FRW) metric and, as a starting point, a flatness assumption. However, we keep the variables associated with the curvature, the reasoning behind which will become clear shortly.

We want to describe the evolution of the Hubble parameter, conformal time and distance measures, all as functions of the logarithmic scale factor, $x=\ln{a}$, working as the main time variable in this paper. This is to be done with the use of the fiducial parameters (``fiducials'') from~\citet{Planckdata}.

The implementation of these functions results in a cosmological model that we can play around with. Our next task is to use observational data from~\citet{supernovadata} to tweak the default cosmological parameters and evaluate their credibility, curvature being one of the parameters. In particular, we will use a Monte Carlo Markov Chain (MCMC) with Metropolis algorithm to explore the parameter space of our model to infer their properties and compare them with observations from supernovae.

MCMC is a popular statistical technique used in various fields besides cosmology. The Metropolis algorithm, a simple and widely used MCMC method, generates a Markov chain of samples in a data set that converge to the target distribution by iteratively accepting or rejecting proposed moves in parameter space based on a set of acceptance criteria.


\subsection{Theory}\label[sec]{mil1:sec:theo}
\input{sections/milestone1/theory.tex}

\subsection{Implementation details}\label[sec]{mil1:sec:imp}
\input{sections/milestone1/implementation.tex}

\subsection{Results}\label[sec]{mil1:sec:res}
\input{sections/milestone1/results.tex}

\subsection{Discussion}\label[sec]{mil1:sec:disc}
\input{sections/milestone1/discussion.tex}