% !TEX root = ./main.tex

\input{overhead.sty}

\begin{document}
% \mainmatter


% \title{Calculating the CMB power spectrum} 
\title{Computing and understanding the CMB anisotropy}
% \title{Microwave music: Finding the cosmic symphony} 
% Microwave Music: Finding the Cosmic Symphony

\author{Nanna Bryne
\inst{1,2}
}
\institute{Institute of Theoretical Astrophysics (ITA), University of Oslo, P.O.~Box 1029 Blindern, N-0315 Oslo, Norway
\and
Center for Computing in Science Education (CCSE), Dept.~of Physics, University of Oslo, P.O.~Box 1048 Blindern, N-0316 Oslo, Norway \\
\email{nanna.bryne@fys.uio.no}}
\titlerunning{Calculating CMB spectrum}
\authorrunning{Bryne} 
\date{\today}  


\abstract{Project in the master-level subject AST5220 Cosmology II at the University of Oslo (UiO).} % context
    {Write a self-contained program to compute theoretical CMB and matter power spectra, starting from fundamental principles.} % aims
    {By perturbing the background and considering the recombination history, we obtained the necessary quantities for calculating the CMB and matter power spectra. To get there, we used the Boltzmann and Einstein equations and linear perturbation theory, amongst other things. We condidered the conformal Newtonian gauge, and we neglected effects from neutrinos, helium, reionisation and polarisation. We used numerical methods s.a.~Runge-Kutta and the trapezoidal rule for integrating.}  % methods
    {Our code provided satisfactory results to the point where neglegcted ingredients s.a. neutrinos were bound to play a role. All results prior to the power spectra essentially substantiate the physics we drew from the latter.} % results
    {We succeded in building a code for our simple \textLambda CDM model with results that can be used to understand the governing physics at large and intermediate scales. However, small-scale physics can be understood better by relatively simple generalisations and improvents to the code.} % conclusions

\keywords{ cosmic microwave background -- %
            large-scale structure of universe}

\maketitle

\bibliographystyle{../../../texmex/aa}

\tableofcontents

\section{Introduction}
\input{sections/introduction}


%%% Milestone I
\section{Background cosmology}\label[sec]{sec:mil1}
\input{sections/milestone1/milestone}


%%% Milestone II
\section{Recombination history}\label[sec]{sec:mil2}
\input{sections/milestone2/milestone}


%%% Milestone III
% \section{Perturbations}\label[sec]{sec:mil3}
\section{Growth of structure}\label[sec]{sec:mil3}
% \section{Evolution of structure in the universe}\label[sec]{sec:mil3}
\input{sections/milestone3/milestone}

%%% Milestone IV
\section{CMB and matter power spectra}\label[sec]{sec:mil4}
\input{sections/milestone4/milestone}

\section{Conclusion}
\input{sections/conclusion}

\begin{acknowledgements}
    Giving thanks to \href{https://github.com/HAWinther/AST5220-Cosmology/}{Hans A. Winther} who is responsible for the skeleton of the code written in this project.% Also thankful for the fact that it is finally over.
\end{acknowledgements}

\bibliography{ref}

\appendix
\input{sections/appendices/appendix.tex}


%\newpage
%\listoffigures


\end{document}