% ----------------------------------
% labels: \label{mil2:[type]:[name]}
% ----------------------------------
% FUTURE TENSE

\colorbox{green}{REPHRASE:} Following the Big Bang (BB), the universe was very hot and dense. Under such extreme conditions, atoms cannot exist. Before atoms formed in the universe, all matter was distributed as a highly ionised plasma. As the universe expanded, the density and temperature decreased. Eventually, the conditions allowed for \textit{recombination} of ions and electrons. That is, atoms (vastly \element{H} and \element{He}) formed during the ``epoch of recombination'' approximately 380\,000 years after BB, at which the temperature had dropped to $\sim3000$~K. Prior to this, the photons were by no means able to travel freely through the ionised plasma as they interacted with free electrons through Thomson scattering. As the free electrons paired up with other baryons to form neutral atoms during recombination, the mean free path (MFP) of the photons increased. The photons emitted at this point in time is what makes up the CMB.

We will in this section examine the free electron fraction and followingly the optical depth and visibility function.

\subsection{Theory}\label[sec]{mil2:sec:theo}
\input{sections/milestone2/theory.tex}

\subsection{Implementation details}\label[sec]{mil2:sec:imp}
\input{sections/milestone2/implementation.tex}

\subsection{Results}\label[sec]{mil2:sec:res}
\input{sections/milestone2/results.tex}

\subsection{Discussion}\label[sec]{mil2:sec:disc}
\input{sections/milestone2/discussion.tex}